---
title: "Course Calendar for Anthropology 4320"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    
    latex_engine: xelatex
  tufte::tufte_book:
    
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r libraries, echo = FALSE}
library(tufte)
library(tidytext)
library(tidyverse)
```

# Welcome to Anthropology 4320: Human Evolution Schedule! 

Dates: 05/26-06/26

Click on a date below to jump to that day!
__________________________________________________________________________________


[Tuesday, May 26 (day 1)]

[Wednesday, May 27 - Decolonizing Human Evolution (day 2)]

[Thursday, May 28 - Evolutionary theory (day 3)]

[Monday, June 1 - Primates (day 4)]

[Tuesday, June 2 -  Bipedalism (day 5)]

[Wednesday, June 3 - The first hominins (day 6)]

[Thursday, June 4 - Early hominin behavior  (day 7)]

[Monday, June  8 - Ardipithecus (day 8)]

[Tuesday, June 9 - Australopithecus! (day 9)]

[Wednesday, June 10 -  Australopithecus continued (day 10)]

[Thursday, June 11 - work catchup (day 11)]

[Monday, June 15 - Early Homo (day 12)]

[Tuesday, June 16 - Homo erectus (day 13)]

[Wednesday, June 17 -  Middle Pleistocene (day 14)]

[Thursday, June 18 - The First *Homo sapiens* (day 15)]

[Monday, June 22  Neandertals - (day 16)]

[Tuesday, June 23  - New species, new ideas (day 17)]

[Wednesday, June 24 - Modern human origins (day 18)]

[Thursday, June 25 - Future (day 19)]



<br />
<br />



# Tuesday, May 26 (day 1)

Today we are going to get things setup for our online class!

**Read the following articles:**

[Read this short blog post](https://beta.capeia.com/paleobiology/2019/09/28/braided-streams-evolutionary-dynamics-among-pleistocene-hominins-in-east-asia-and-the-evolution-of-homo-sapiens)
[and this interview](https://gizmodo.com/this-philosopher-is-challenging-all-of-evolutionary-psy-1842248835)


### Create an account on Perusall (make sure to use the class code sent via email when you sign up so you get access to our site ) 


### Join the class Slack

- You will get an email from me with info on how to join!
- you can use Slack on your phone or on a computer. I tend to use the phone but either works
- post a Hello message in the say-hello channel to get a handle on how it works. I'll be on Slack all day to help folks set things up


### Create a blog

- [Click here for more information and for an overview of how to set this up](https://marckissel.github.io/Anthro_4320/blog_notes.html)

 **- When you have created your blog send me a link!**


<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />


# Wednesday, May 27 - Decolonizing Human Evolution (day 2)

Today we are focusing on a issue that isn't always addressed but is of critical importance: Who gets to tell the story of human origins. All 3 of these papers discuss important aspects about how feminist and anti-racist scholars have re-framed the discussions of human evolution. Athreya and Ackermann look at the role colonialism played in the founding of the discipline. Dunsworth asks us to think about how we teach these topics in class. And Sterling look as how Black feminist theory can address the
sociopolitics of archaeological practice. 

<br />
<br />

**Read the following articles:**

Note: All the readings are on AsULearn and Persuall. If i was able to find open access versions of the papers i linked them here just to make it easier, but this wasn't always possible :(

- [Athreya and Ackermann 2018 Colonialism Narratives Human Origins Asia Africa](https://osf.io/f43tr/)

- [Dunsworth_2018_It is unethical to teach evolution without confronting racism and sexism](https://evolution-institute.org/it-is-unethical-to-teach-evolution-without-confronting-racism-and-sexism/) 

- [Sterling_2015_Black Feminist Theory in Prehistory](https://link.springer.com/article/10.1007/s11759-015-9265-z)


```{marginfigure, echo= TRUE}


**Here's why I Want you to do it:**

We want to begin to think about who gets to tell the story of human origins. As we all know, there is a lot of controversy about where we come from. But  rather than center a debate about creationism (though you are free to explore that in your post!) I thought it would be interesting to see how recent work  is pushing against the narrative in a different way 


**Here's how to do it:**
  
Use the 3 papers (and any others you know of/can find) as a scaffold for the post. You don't have to reference them all and there is no set format for referencing
Think about [adding visuals to your blog](https://oudigitools.blogspot.com/2020/05/images-in-posts.html)

```

**Now, Here's what I want you to do:**

1. Write your first blog post based on the readings.  One topic could be how your previous classes have dealt w/ the issues raised in the readings; examples of how this can be done better; or simply what you thought while reading the papers. You could also use the readings as a jumping off point for disucssing other aspects of anthropology that intersect with these ideas 


***Please have this posted by Friday May 29 at 5:00 pm if possible***



<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />





# Thursday, May 28 - Evolutionary theory (day 3)
Today we want to think about what we know about evolutionary theory. There is only one reading but it is both fairly long and a classic of evolutionary thought. Think about what Gould & Lewontin are saying and how their framing the processes of evolution. To give some context, this paper has been cited over 8900 times! 
The video then gives some context on how paleoanthropolgists see the species concept.

<br />
<br />

**Read the following articles:**

- [Gould and Lewontin 1979 The Spandrels of San Marco and the Panglossian Paradigm-A Critique of the Adaptationist Programme](https://royalsocietypublishing.org/doi/pdf/10.1098/rspb.1979.0086)


watch this video:
[Species Concepts in Palaeoanthropology](https://www.youtube.com/watch?v=GWmtak1BGxo)

**Now**

- make annotations on the  article on the Persuall site. You could ask for clarification on these ideas or just explore your thoughts

- Optional Zoom meeting: 

If you can make it we will jump on to Zoom and Slack to talk about the articles for this week (will post a poll for times). Let me know if you want to come and we can setup a Zoom link!

<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Weekend work 

1. Pick a fossil from the list online () and begin to learn about it. Jot down notes etc
2. Follow the directions [here to start with Aris...](https://marckissel.github.io/Anthro_4320/online_game_notes.html) and work though the tutorial.

<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Monday, June 1 - Primates (day 4)

This class day is set to remind us what we learned in previous Bioanth classes! Some background on primates and early hominin evolution. Feel free to skim the chpt on Early Hominins if you remember them!
<br />
<br />

**Read the following articles:**

Primate Cognition[https://www.nature.com/scitable/knowledge/library/primate-cognition-59751723/]

[Early Hominins](http://explorations.americananthro.org/wp-content/uploads/2020/02/Chapter-9-Early-Hominins3.0.pdf)


```{marginfigure, echo= TRUE}


**Here's why I Want you to do it:**

This is not a dumb question nor is it easy to answer. I want you to think about how you could respond to this question in a way that is respectful and useful. We will return to this question at the end of the class as well


**Here's how to do it:**

Using your knowledge from other classes, as well as any other sources, write a conversational response and share it on your blog. If you want, feel free to insert a video, gif, or image. You may want to look at examples of how to engage with the public and how to craft responses that are useful.

```

**Now, Here's what i want you to do:**

Write your blog post based on answering the following question:

>
> If humans came from chimps, why are there still chimps?
>

***Please post this by Thursday, June 4th***

<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Tuesday, June 2 -  Bipedalism (day 5)

Today we are focusing on bipedalism. Remember that one of the ways we define a hominin is by it walking upright on two legs. The piece by Lovejoy is sort of a classic in paleoanthropology.This doesn't mean it is correct. It has however generated a lot of discussion. The second piece is a bit long but it has a nice section that talks about different theories of the origins of bipedality, so make sure to at least look at the section on "Models of human bipedal acquisition"

Use the Persuall site to mark up and comment on what you read!

<br />
<br />
**Read the following articles:**


[Lovejoy 1981-The origin of man](https://www.researchgate.net/publication/6092674_The_Origin_of_Man) 
[Kimura_2019_How did humans acquire erect bipedal walking](https://www.jstage.jst.go.jp/article/ase/127/1/127_190219/_pdf)


**watch this video**

[When We First Walked](https://www.youtube.com/watch?v=3bFtotU0of4)


Read this if you want some more detailed info: [Ward et al 2019 A Late Miocene Hominid Partial Pelvis From Hungary](https://www.ncbi.nlm.nih.gov/pubmed/31537304)



**Now**, 

- Make comments on Persual on these papers. No set number of comments needed but engage at the level you feel is needed. Feel free to use info from the videos or other readings to inform your post

- Work on the Sutori site for your fossil. If you want comments send me a message and i will tak a look at it ASAP

- Think about how you want your section of the Aris game to look. Plan it out on paper so you have a gameplan and share any concerns on the Slack channel

<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />


# Wednesday, June 3 - The first hominins (day 6)

One of the most contentious issues in human evolution is who the first hominid was. Today we are going to look at 2 plausible candidates, *Orrorin tugenensis* and *Sahelanthropus tchadensis*. These papers are very technical but they give a good understanding of what we know about the species. Feel free to explore more and see what else you can find
<br />
<br />
if you want a refresher on these fossils here is a short post i wrote on each:
[*Sahelanthropus tchadensis*](https://marckissel.netlify.app/post/day1-sahelanthropus-tchadensis/)
[*Orrorin tugenensis*](https://marckissel.netlify.app/post/day2-orrorin-tugenensis/)

**Read the following article:**

[Brunet et al 2002 - A new hominid from the Upper Miocene of Chad, Central Africa](http://planet.botany.uwc.ac.za/nisl/research_methods/scientific_methods/pdf/chad_hominids.pdf)
[Pickford et al - Bipedalism in Orrorin tugenensis revealed by its femora](https://www.sciencedirect.com/science/article/pii/S1631068302000283) (n.b: feel free to skip the French translation)

+++ if you want more info, this piece on the [Orroin femur could help](https://cashp.columbian.gwu.edu/sites/g/files/zaxdzs1746/f/downloads/2008-Richmond-Jungers.pdf)

+++ and here is a fascinating [discussion of the femur of Sahelanthropus](https://www.nature.com/articles/d41586-018-00972-z)

```{marginfigure, echo= TRUE}


**Here's why I Want you to do it:**
Tracing the fossil evidence is difficult and this literature is really hard to parse. By starting to keep track of something like this you can begin to build your knowledge


**Here's how to do it:**
look at one of the articles over the past 2 days and notice what they are talking about in terms of functional aspects of bipdealism. For example, what about the femur of *Orroin* suggests it is a biped. If you need help [this site is full of awesome info](https://humanorigins.si.edu/human-characteristics/walking-upright). Or you could decide to look at what make the foramen magnum relevant for bipedalism. If you need help, send me a message on Slack! 
```

**Now, here's what i want you to do** 

- Pick one aspect of the skeletal evidence of bipedalism and write a short blog post about it. If you need more background check out the chapter on early hominins from Monday or do a web search. 



<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Thursday, June 4 - Early hominin behavior  (day 7)

Taking a break from reading (yay!) we are going to watch some lectures from leaders in the field. The first, by Carol Ward, talks about early hominin body form. The second looks at the role scavenging and hunting played in human history. And the third asks about how aggression and warfare may have been important pressures in human origins. 

we can jump on Slack to watch together at a certain time [will post time soon] or watch them by yourself and comment on slack afterwards. The readings from the previous class should help inform on these videos.

<br />
<br />


[Early Hominin Body Form](https://carta.anthropogeny.org/mediaplayer/play/5367/3264)
[The Ecology of Hominin Scavenging](https://carta.anthropogeny.org/mediaplayer/play/313788/302508)
[The Parallel Evolution of Humanity and Savagery](https://carta.anthropogeny.org/mediaplayer/play/20547/7939)



<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Weekend

- Continue on planning out your part of the game. You already have a fossil chosen so begin to find articles on it and read them. If you have time, finish your Sutori site 

<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Monday, June  8 - Ardipithecus (day 8)
The story about Ardipithecus is sort of almost mythical in human evolution. Today we are trying something different. We are all going to watch a short video and then read one overview piece about the species. Then, you are going to choose one of the more in-depth articles and read that carefully. Use that paper to write up a short blog post.
<br />
<br />

**Watch this short video**:

[The Analysis of Ardipithecus ramidus -- One of the Earliest Known Hominids](https://www.youtube.com/watch?v=EC9aIth1ah4&)



**Read the following articles:**

[White et al. 2009 Ardipithecus ramidus and the Paleobiology of Early](https://science.sciencemag.org/content/326/5949/64)









```{marginfigure, echo= TRUE}


**Here's why I Want you to do it:**
Arid was published 11 years ago. Since then a lot has changed but we still don't know that much about this time period. I want you to focus on one aspect of its anatomy. You can then 'follow' that part as we continue in our understanding of human origins


**Here's how to do it:**
As you read the article, take notes and ask questions (on Persuall might help for feedback!). Get the larger picture of what is going on in the form and function. if there is a bone, element, whatever that you don't know look it up online or ask in the Slack.  
```


**Now, Here's what i want you to do:**


Pick one of the research articles below and read it in depth. 
[links to articles](https://www.sciencemag.org/site/feature/misc/webfeat/ardipithecus/index.xhtml). 

Prepare a blog post for *Friday June 12th* on the article you read. You should do your best to summarize the article as well as talk about its significance and what you learned. You could also go further and see what recent data exists that may update what we know (note: you might not find anything new, in which case that is also something important to note)


<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />





# Tuesday, June 9 - Australopithecus! (day 9)

While traditionally known as "bipedal apes" Australopithecines are now seen in a different light. Some scholars think they even may have made and used stone tools! The piece by Ward and Hammond covers a lot of useful ground and is a good starting point. The Ward et al. 1999 is more technical and you can skim that if you wish (but full of interesting stuff!). The Halie-Selassie et al. paper is a newer article that talks about the face of anamensis (I also included a news article that covers the main points of that paper)

There is also an optional Zoom hangout today at 1 (time may change!). Stop by to chat about the class, the papers, or to see how my kids love to interrupt me while I'm working

<br />
<br />
**Read the following article:**

[Ward and Hammond 2016 Australopithecus and Kin](https://www.nature.com/scitable/knowledge/library/australopithecus-and-kin-145077614/)

[Ward et al. 1999 The new hominid species Australopithecus anamensis](https://onlinelibrary.wiley.com/doi/abs/10.1002/(SICI)1520-6505(1999)7:6%3C197::AID-EVAN4%3E3.0.CO;2-T?casa_token=KoYvaznX7fwAAAAA:cU_dwFzG9U7oNHvQtboGt5QyINEk1qBYSIVbI7WutYq-_hSIojDh6fDbNoW_Li_eN7WVB37TyUV3tXr4)

[Halie-Selassie et al 2019. A 3.8-million-year-old hominin cranium from Woranso-Mille, Ethiopia](https://www.nature.com/articles/s41586-019-1513-8)

[Here is a news article that summarisze this work](https://www.sciencenews.org/article/australopithecus-anamensis-skull-reveals-face-lucy-possible-ancestors)

**Watch this video**

[Visiting the Taung Child at Wits University](https://www.youtube.com/watch?v=0mWOJ56_Et0)

**Now**, 

1. Continue work on the blog post from yesterday. Incorperate, if possible, information from these papers that link Arid to the first australopiths

2. Optional Zoom meeting: 



<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />


# Wednesday, June 10 -  Australopithecus continued (day 10)

Since we can't cover all the info on Australopiths(i took a grad seminar that solely focused on them and even then we ran out of time!) today we are going to learn about one of the more recently discovered one, *Australopithecus sediba*.  The first paper by Berger and colleagues introduces the species and the second one gives a more skeptical look. Then watch the video for a new take on a different Australopith and its behavior!



**Read the following article:**

[Berger et al. 2010. Australopithecus sediba: A New Species of Homo-Like Australopith from South Africa](https://www.academia.edu/1586502/Australopithecus_sediba_A_New_Species_of_Homo-Like_Australopith_from_South_Africa)
[Kimble 2013](Hesitation on hominin history)

If you want more information, check out some of the [papers in this special volume devoted to *Au. sediba*](http://www.paleoanthro.org/journal/volumes/2018/)

**then watch this video**

[Thumb can tell if Paranthropus boisei made stone tools](https://www.youtube.com/watch?v=cEkpNbjUcfk)

**Now**, 

1. Work on Aris game

2. Post thoughts/comments on Perusall


<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Thursday, June 11 - work catchup (day 11) 

Today we are taking a break (phew!) and catching up on class projects. Work on all the things. I'll be on Zoom for most of the day if you want to chat! 

**no assigned reading**



<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />


# Weekend work 

ADD

<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Monday, June 15 - Early Homo (day 12)

One thing we don't know well is when the genus *Homo* first evolved. The Anton et al paper is one of my favorite takes on this question and it covers a lot of ground. Make some annotations on it in Persuall or send a question to the group Slack if you are confused by something
 


**Read the following article:** 

[Anton et al. 2014 Evolution of early Homo: An integrated biological perspective](https://science.sciencemag.org/content/345/6192/1236828)



**Now, here's what I want you to do**,

Post comments on the Anton paper on the Persuall site.


<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Tuesday, June 16 - Homo erectus (day 13) 

Today we are looking at two papers that came out in April! I haven't had time to really think about them yet so excited to see what these papers have to say, but they have to do with the question of how many hominins were around at the same time

**Read the following articles:**

[Semaw el al 2020. Co-occurrence of Acheulian and Oldowan artifacts with Homo erectus cranial fossils from Gona, Afar, Ethiopia](https://advances.sciencemag.org/lens/advances/6/10/eaaw4694)

[Herries et al. 2020 Contemporaneity of Australopithecus, Paranthropus, and early Homo erectus in South Africa](https://www.uj.ac.za/newandevents/Pages/Contemporaneity-of-Australopithecus,-Paranthropus,-and-early-Homo-erectus-in-South-Africa.aspx)


**then, watch this video**

[One cave. Three hominin lineages. Australopithecus and Paranthropus meet Homo in Drimolen](https://www.youtube.com/watch?v=MOJrfa8Flus)

Bonus (in case you want to know more about Homo erectus:

[Anton et al. 2016 Morphological variation in Homo erectus and the origins of developmental plasticity](https://royalsocietypublishing.org/doi/10.1098/rstb.2015.0236)


```{marginfigure, echo= TRUE}


**Here's why I Want you to do it:**
We don't know what this would have been like, but it seems clear that multiple hominins were around at the same time. How would we treat a different species in the same Family today?


**Here's how to do it:**
some other things to look at would be info on *Homo floresiensis*
```

**Now, here's what I want you to do**

- This is a different sort of blog post. Playing off the readings for today and your knowledge of anthropology I want you to think about what it would have been like for two hominin species to meet. Uou could do this as a sort of narrative piece, you could make up a story, etc. I don't have a set form in mind so take it in any direction you like.





<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Wednesday, June 17 -  Middle Pleistocene (day 14)

The Middle Pleistocene (sometimes called the *Muddle in the Middle*) is a confusing time period. Trying to cover it is even harder. I picked two articles here that I think might be interesting. The first, by Ackermann &  colleagues, is a very important paper that talks about where our species comes from. The second one is also interesting and demonstrates some of the latest ideas coming out. The third is a nice overview that just came out this week!

[This textbook chapter by Paskey & Cisneros gives some great background if you are rusty on these species](http://explorations.americananthro.org/wp-content/uploads/2020/01/Chapter-11-Archaic-Homo-2.0.pdf)

**Read the following:**

[Ackermann et al. 2015 The Hybrid Origin of ‘‘Modern’’ Humans](https://link.springer.com/article/10.1007/s11692-015-9348-1)

[Communicative capacities in Middle Pleistocene humans from the Sierra de Atapuerca in Spain](https://www.researchgate.net/publication/8495575_Auditory_capacities_in_Middle_Pleistocene_humans_from_the_Sierra_de_Atapuerca_in_Spain/link/0912f50fa6157da75d000000/download) NEED
                                                                                                        
[When Did Homo Sapiens First Appear?](https://www.discovermagazine.com/planet-earth/when-did-homo-sapiens-first-appear)

**Now**, 
- make sure your part of Aris is ready!

- Optional Zoom meeting to talk about class projects and readings


<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />


# Thursday, June 18 - The First *Homo sapiens* (day 15)

Today we are looking at what it means to be in the species *Homo sapiens*. The two papers are a bit technical (though Stringer's is easier) but they cover a lot of ground. Think about how we define our species and what pushing back the date of the first appearance of our species means.


[Hublin et al. 2017](New fossils from Jebel Irhoud, Morocco and the pan-African origin of Homo sapiens)

[Stringer 2016 The origin and evolution of Homo sapiens](https://royalsocietypublishing.org/doi/full/10.1098/rstb.2015.0237)

**Now here's what I want you to do**,

Write one last blog post. This one should focus on answering the following question:

On a visit to D.C. (post-pandemic, of course), you happen to find yourself standing in an elevator next to congressional leaders taking a tour of the museum you are in. They notice you are carrying some papers about human evolution and they ask you what they are about. With that opening, lay out your best four minute ‘elevator speech’ as to why the next federal budget should include additional funding for the study of human evolution

Please post by  Wednesday, June 24

<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Weekend work 


<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Monday, June 22  Neandertals - (day 16)

Our department teaches a whole class on Neandertals so it is of course hard to cover them in a day. I posted here two new papers on the Neandertal genome

**Read the following article:**

[Skov et al. 2020 The nature of Neanderthal introgression revealed by 27,566 Icelandic genomes](https://www.nature.com/articles/s41586-020-2225-9)
[Taskent et al 2020. Analysis of Haplotypic Variation and Deletion Polymorphisms Point to Multiple Archaic Introgression Events, Including from Altai Neanderthal Lineage](https://www.genetics.org/content/genetics/early/2020/03/31/genetics.120.303167.full.pdf)



<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Tuesday, June 23  - New species, new ideas (day 17) 

In the last few years our knowledge of the fossil record of human evolution has greatly expanded. These 2 articles talk about newly proposed species. Read them carefully and think about how they are both changing the landscape of human origins

**Read the following article:**

[Berger et al 2017 Homo naledi and Pleistocene hominin evolution in subequatorial Africa](https://elifesciences.org/articles/24234)
[A new species of Homo from the Late Pleistocene of the Philippines](https://www.nature.com/articles/s41586-019-1067-9)

**Now**, 

We are going to watch this documentary together:
https://www.youtube.com/watch?v=RzLJAa5X4Fo&t=2s


<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Wednesday, June 24 - Modern human origins (day 18) 

Ok, but why are *Homo sapiens* the only hominins left on the planet. This is another hot topic in paleoanthropology. The 

**Read the following article:**

[Scerri et al. 2018 Did our species evolve in subdivided populations across Africa, and why does it matter?](https://www.researchgate.net/publication/326408394_Did_Our_Species_Evolve_in_Subdivided_Populations_across_Africa_and_Why_Does_It_Matter)
[Hublin et al. 2020 Initial Upper Palaeolithic Homo sapiens from Bacho Kiro Cave, Bulgaria](https://www.nature.com/articles/s41586-020-2259-z)



**Now**, 





<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />
<br />

# Thursday, June 25 - Future (day 19) 

What does the future hold?

**Read the following article:**

[Prehistoric human footprints reveal a rare snapshot of ancient human group behavior](https://theconversation.com/prehistoric-human-footprints-reveal-a-rare-snapshot-of-ancient-human-group-behavior-138502)


[Expanding the evolutionary explanations for sex differences in the human skeleton](https://onlinelibrary.wiley.com/doi/epdf/10.1002/evan.21834)

**Now**, 

Zoom meeting

# Friday, June 26  - wrapup! (day 20)

Last day!

Test game















.




.

